{\small} High Performance P$^{3}$M N-body code: CUBEP3M:

### 2012-05-31 Halos.tex
Incorporated Hughs comments, plus a catastrophic plot showing what happens if we turn off the random offset in the current code

Once the list of peak positions is generated, they are sorted from the highest 
to the lowest density value. Then each of the halo candidates is inspected 
independently, starting with the highest peak. The grid mass is accumulated 
in spherical shells of fine grid cells surrounding the maximum, until the 
mean density within the halo drops below a pre-defined overdensity cutoff 
(usually set to 178 in units of the mean, in accordance to the top-hat 
collapse model). As we accumulate mass we remove it from the mesh, so that no 
mass element is double-counted. This method is thus inappropriate for finding 
sub-halos as within this framework those are naturally incorporated in their 
host halos. Because the halos are found on a grid of finite-size cells, it is 
possible, especially for the low-mass halos, to overshoot the target overdensity.
When this occurs we use an analytical halo density profile to correct the 
halo mass and radius to the values corresponding to the target overdensity. 
This analytical density profile is given by Truncated Isothermal Sphere (TIS) 
profile \citep{1999MNRAS.307..203S,2001MNRAS.325..468I} for overdensities below 
$\sim130$, and $1/r^2$ for lower overdensities. The TIS density profile has a
similar outer slope (the relevant one here) to the Navarro, Frenk and White 
\citep[NFW][]{1997ApJ...490..493N} profile, but extends to lower overdensities
and matches well the virialization shock position given by the Bertschinger 
self-similar collapse solution \citep{1985ApJS...58...39B}.

@@

A sample halo mass function produced based on our inlined SO halo finder 
at redshift $z=0$ from a RANGER4000 simulation is shown in Fig. \ref{mf}. We compare our result to the 
precise fit presented recently by \citet{2008ApJ...688..709T}. Unlike most
other widely-used fits like the one by \citet{2002MNRAS.329...61S}, which are based on friends-of-friends (FOF)
halo finders, the \citet{2008ApJ...688..709T} fit is based on the
SO search algorithm, whose masses are systematically different 
from the FOF masses \citep[e.g.][]{2007MNRAS.374....2R,2008ApJ...688..709T}, 
making this fit a better base for comparison here. Results show excellent
agreement, within $\sim10$ per cent for all halos with masses corresponding to
1000 particles or more. Lower-mass halos are somewhat under-counted compared
to the \citet{2008ApJ...688..709T} fit, by $\sim20$ per cent for 400 particles and 
by $\sim40$ per cent for 50 particles. This is due to the grid-based nature of our
SO halo finder, which misses some of the low-mass halos. Using more sophisticated
halo finders (available through post-processing due to their heavier memory
footprint) recovers the expected mass function.

{\bf II: do we need to say anything more about having the option to use other
halo finders like AHF and FOF?}

A second test of the accuracy of the halo finder algorithm is to extract the halo power spectrum $P_{h}(k)$ and compare the halo bias
with theoretical predictions. The halo bias is defined as $b(k) = \sqrt{P(k)/P_{h}(k)}$ and is shown in Fig. \ref{fig:halo}.

\begin{figure}%[ht]
%  \vskip -0.5cm 
  \begin{center}
    \includegraphics[width=3.2in]{graphs/bias.eps}
  \end{center}
  \caption{Halo bias, measured at $z=0$ from a $4000^3$ particles simulations with a side length of $3.2 h^{-1}$Gpc.
  Bottom to top curves correspond to haloes of different masses, from light to massive. The four bins are decades
  of halo masses; the first contains haloes made of 20-200 particles, the second of 200-2,000 and so on.
    \label{fig:halo}}
\end{figure}

>>>>

Once the list of peak positions is generated, they are sorted from the highest 
to the lowest density value. Then each of the halo candidates is inspected 
independently, starting with the highest peak. The grid mass is accumulated 
in spherical shells of fine grid cells surrounding the maximum, until the 
mean density within the halo drops below a pre-defined overdensity cutoff 
(usually set to 178 in units of the mean, in accordance to the top-hat 
collapse model). As we accumulate mass we remove it from the mesh, so that no 
mass element is double-counted. This method is thus inappropriate for finding 
sub-halos as within this framework those are naturally incorporated in their 
host halos. Because the halos are found on a grid of finite-size cells, it is 
possible, especially for the low-mass halos, to overshoot the target overdensity.
When this occurs we use an analytical halo density profile to correct the 
halo mass and radius to the values corresponding to the target overdensity. 
This analytical density profile is given by Truncated Isothermal Sphere (TIS) 
profile \citep{1999MNRAS.307..203S,2001MNRAS.325..468I} for overdensities below 
$\sim130$, and $1/r^2$ for lower overdensities. The density profile has a
similar outer slope (the relevant one here) to the Navarro, Frenk and White 
\citep[NFW][]{1997ApJ...490..493N} profile, but extends to lower overdensities
and matches well the virialization shock position given by the Bertschinger 
self-similar collapse solution \citep{1985ApJS...58...39B}.

@@

A sample halo mass function produced based on our inlined SO halo finder 
at redshift $z=0$ from a RANGER4000 simulation is shown in Fig. \ref{mf}. We compare our result to the 
precise fit presented recently by \citet{2008ApJ...688..709T}. Unlike most
other widely-used fits like the one by \citet{2002MNRAS.329...61S}, which are based on friends-of-friends (FOF)
halo finders, the \citet{2008ApJ...688..709T} fit is based on the
SO search algorithm, whose masses are systematically different 
from the FOF masses \citep[e.g.][]{2007MNRAS.374....2R,2008ApJ...688..709T}, 
making this fit a better base for comparison here. Results show excellent
agreement, within $\sim10$ per cent for all halos with masses corresponding to
1000 particles or more. Lower-mass halos are somewhat under-counted compared
to the \citet{2008ApJ...688..709T} fit, by $\sim20$ per cent for 400 particles and 
by $\sim40$ per cent for 50 particles. This is due to the grid-based nature of our
SO halo finder, which misses some of the low-mass halos. It was shown, however, that using more sophisticated
halo finders (available through post-processing due to their heavier memory
footprint) allows to recover the expected mass function.

A second test of the accuracy of the halo finder algorithm is to extract the halo power spectrum $P_{h}(k)$ and compare the halo bias
with theoretical predictions. The halo bias is defined as $b(k) = \sqrt{P(k)/P_{h}(k)}$ and is shown in Fig. \ref{fig:halo}.
{\bf Expand here, describe, match with theory, etc...}
\begin{figure}%[ht]
%  \vskip -0.5cm 
  \begin{center}
    \includegraphics[width=3.2in]{graphs/bias.eps}
  \end{center}
  \caption{Halo bias, measured at $z=0$ from a $4000^3$ particles simulations with a side length of $3.2 h^{-1}$Gpc.
  Bottom to top curves correspond to haloes of different masses, from light to massive. The four bins are decades
  of halo masses; the first contains haloes made of 20-200 particles, the second of 200-2,000 and so on.
    \label{fig:halo}}
\end{figure}

### 2012-05-31 Introduction.tex
Incorporated Hughs comments, plus a catastrophic plot showing what happens if we turn off the random offset in the current code

The basic problem that is addressed with N-body codes is a time evolution of an ensemble of $N$ particles
that is subject to gravitational attraction. The brute force calculation requires $O(N^{2})$ operation, a cost that 
exceeds the memory and speed of current machines for large problems.
Solving the problem  on a mesh \citep{1981csup.book.....H} reduces to $O(N\mbox{log}N)$ the number of operations,
as it is possible to solve for the particle-mesh (PM) interaction with fast Fourier transforms techniques, 
typically using high performance libraries such as {\small FFTW} \citep{FFTW3}.

With the advent of large computing facilities, parallel coding has now become 
common practice, and N-body codes have evolved both in performance and complexity. 
Many codes have opted for a tree algorithm, including {\small GADGET} \citep{2001NewA....6...79S, 2005MNRAS.364.1105S}, PMT \citep{1995ApJS...98..355X}, GOTPM \citep{2004NewA....9..111D}, and Hydra \citep{1995ApJ...452..797C}, in which the local resolution increases with the density of the matter field. 
These have the advantage to balance the load across the computing units, which enable the calculation of high density regions. 
The drawback is a significant loss in speed, which can be only partly recovered by turning off the tree algorithm. 
The same reasoning applies to the mesh-refine code of  \cite{1991ApJ...368L..23C} and the PM code by \cite{1995astro.ph..3042F}.
 these are not designed to perform large scale PM calculations.

{\bf ( PM code by japanese? Others codes I should mention?)}

{\small PMFAST} (\cite{2005NewA...10..393M}, MPT hereafter) is one of the first code that is designed such as to optimize the PM algorithm,
both in terms of speed and memory usage. It uses a two-level mesh algorithm based on the gravity solver of \cite{2003AAS...203.9703T},
The long range gravitational force is computed on a  grid four times coarser, such as to minimize the {\small MPI} communication time
and to fit in system's memory. The short range is computed locally on a finer mesh, and only the local sub-volume needs 
to be stored at a given time, allowing for {\small OPENMP} parallel computation.
This this setup enable the code to evolve large cosmological systems both rapidly and accurately, on relatively modest clusters.
One of the main advantage of {\small PMFAST} over other PM codes is that the number of large arrays is minimized,
and the global {\small MPI} communications are cut down to the minimum: for passing particle at the beginning of each time step,
and  for computing the long range FFTs.
As described in MPT, access to particles is accelerated with the use of linked lists, deletion of `ghost' particles
in buffer zones is done at the same time as particles are passed to adjacent nodes,
and the global FFTs are performed with a slab decomposition of the volumes via a special file transfer interface, 
designed specifically to preserve a high processor load.

Since its first published version, {\small PMFAST} has evolved in many aspects. 
The first major improvement was to transform the volume decomposition in multi-node configurations 
from slabs to cubes. One of the problem with slabs is that they do not scale well to large runs: 
as the number of cells per dimension increases, the thickness of each slab shrinks rapidly,
until it reaches the hard limit of a single cell layer.  With this enhancement, the code name was changed to {\small CUBEPM},
which soon after incorporated particle-particle (pp) interactions at the sub-grid level. 
The code was finally renamed {\small CUBEP3M}, and now includes a significant number of new features: the pp force
can be extended to an arbitrary range, the size of the redshift jump can be constrained for improved accuracy during the first time steps,
a runtime halo finder has been implemented, the expansion has also been generalized to include a redshift dependent equation of state of dark energy, there is a system of unique particle identification which can be switched on or off, and the initial condition generator has been generalized as to include non-Gaussian features {\bf (anything else since PMFAST?)}.

This paper aims at presenting and quantifying these new features that make {\small CUBEP3M} one of the most competitive public N-body code.
It has already been involved in a number of scientific applications over the last few years,
spanning the field of weak lensing \citep{Vafaei10, 2008MNRAS.388.1819L,  2009arXiv0905.0501D, 2010PhRvD..81l3015L, 2010arXiv1012.0444Y, 2012arXiv1202.2332H},  BAO
 (\cite{2010arXiv1008.3506Z,  2011arXiv1106.5548N, 2011arXiv1109.5746H}; Harnois-D\'{eraps} et al 2012 ({\bf update with last paper!})), 
formation of early cosmic structures \citep{2008arXiv0806.2887I,2010arXiv1005.2502I},
observations of dark stars \citep{2010MNRAS.407L..74Z,2012MNRAS.tmp.2794I} and
reionization \citep{2011arXiv1107.4772I,Fernandez:2011ab,2011MNRAS.413.1353F,2012MNRAS.422..926M,Datta:2011hv,2012arXiv1203.0517F},  
and it is thus important for the community to have access to a paper that describes the methodology, the accuracy and the performance of this public code.

>>>>

The basic problem that is addressed with N-body codes is the time evolution of an ensemble of $N$ particles
that is subject to gravitational attraction. The brute force calculation requires $O(N^{2})$ operations, a cost that 
exceeds the memory and speed of current machines for large problems.
Solving the problem  on a mesh \citep{1981csup.book.....H} reduces to $O(N\mbox{log}N)$ the number of operations,
as it is possible to solve for the particle-mesh (PM) interaction with fast Fourier transforms techniques with high performance libraries such as {\small FFTW} \citep{FFTW3}.

With the advent of large computing facilities, parallel computations have now become 
common practice, and N-body codes have evolved both in performance and complexity. 
Many have opted for  `tree' algorithms, including {\small GADGET} \citep{2001NewA....6...79S, 2005MNRAS.364.1105S}, {\small PMT} \citep{1995ApJS...98..355X}, {\small GOTPM} \citep{2004NewA....9..111D}, and Hydra \citep{1995ApJ...452..797C}, in which the local resolution increases with the density of the matter field. 
These often have the advantage to balance the work load across the computing units, which enable fast calculations even in high density regions. 
The drawback is a significant loss in speed, which can be only partly recovered by turning off the tree algorithm. 
The same reasoning applies to mesh-refined codes  \citep{1991ApJ...368L..23C}, 
which in the end are not designed to perform fast PM calculations on large scales.

%{\bf ( PM code by japanese? Others codes I should mention?)}

{\small PMFAST} (\cite{2005NewA...10..393M}, MPT hereafter) is one of the first code designed specifically to optimize the PM algorithm,
both in terms of speed and memory usage. It uses a two-level mesh algorithm based on the gravity solver of \cite{2003AAS...203.9703T},
The long range gravitational force is computed on a  grid four times coarser, such as to minimize the {\small MPI} communication time
and to fit in system's memory. The short range is computed locally on a finer mesh, and only the local sub-volume needs 
to be stored at a given time, allowing for {\small OPENMP} parallel computation.
This this setup enable the code to evolve large cosmological systems both rapidly and accurately, on relatively modest clusters.
One of the main advantage of {\small PMFAST} over other PM codes is that the number of large arrays is minimized,
and the global {\small MPI} communications are cut down to the minimum: for passing particle at the beginning of each time step,
and  for computing the long range FFTs.
As described in MPT, access to particles is accelerated with the use of linked lists, deletion of `ghost' particles
in buffer zones is done at the same time as particles are passed to adjacent nodes,
and the global FFTs are performed with a slab decomposition of the volumes via a special file transfer interface, 
designed specifically to preserve a high processor load.

Since its first published version, {\small PMFAST} has evolved in many aspects. 
The first major improvement was to transform the volume decomposition in multi-node configurations 
from slabs to cubes. One of the problem with slabs is that they do not scale well to large runs: 
as the number of cells per dimension increases, the thickness of each slab shrinks rapidly,
until it reaches the hard limit of a single cell layer.  With this enhancement, the code name was changed to {\small CUBEPM},
which soon after incorporated particle-particle (pp) interactions at the sub-grid level. 
The code was finally renamed {\small CUBEP3M}, and now includes a significant number of new features: the pp force
can be extended to an arbitrary range, the size of the redshift jump can be constrained for improved accuracy during the first time steps,
a runtime halo finder has been implemented, the expansion has also been generalized to include a redshift dependent equation of state of dark energy, there is a system of unique particle identification which can be switched on or off, and the initial condition generator has been generalized as to include non-Gaussian features.
 {\small PMFAST} was also equipped with a multi time stepping option that has not been tested on {\small CUBEP3M} yet, but which is, in principle at least, still available. 
 It also contains support for gas cosmological evolution through a TVD MHD module,
 and a coupling interface with the radiative transfer code C2-Ray \citep{C2-Ray...}.

This paper aims at presenting and quantifying these new features that make {\small CUBEP3M} one of the most competitive public N-body code.
It has already been involved in a number of scientific applications over the last few years,
spanning the field of weak lensing \citep{Vafaei10, 2008MNRAS.388.1819L,  2009arXiv0905.0501D, 2010PhRvD..81l3015L, 2010arXiv1012.0444Y, 2012arXiv1202.2332H},  BAO
 \citep{2010arXiv1008.3506Z,  2011arXiv1106.5548N, 2011arXiv1109.5746H, 2012arXiv1205.4989H}, 
formation of early cosmic structures \citep{2008arXiv0806.2887I,2010arXiv1005.2502I},
observations of dark stars \citep{2010MNRAS.407L..74Z,2012MNRAS.tmp.2794I} and
reionization \citep{2011arXiv1107.4772I,Fernandez:2011ab,2011MNRAS.413.1353F,2012MNRAS.422..926M,Datta:2011hv,2012arXiv1203.0517F},  
and it is thus important for the community to have access to a paper that describes the methodology, the accuracy and the performance of this public code.

### 2012-05-31 PoissonSolver.tex
Incorporated Hughs comments, plus a catastrophic plot showing what happens if we turn off the random offset in the current code

This section reviews how Poisson's equation is solved on a double-mesh configuration. 
Many parts of the algorithm are identical to {\small PMFAST}, hence we refer the reader 
to section 2 of MPT for more details. In {\small CUBEP3M}, the mass default assignment scheme are
a `cloud-in-cell' (cic) interpolation for the coarse grid,  and a `nearest-grid-point' interpolation 
for the fine grid \citep{1981csup.book.....H}. Particle-particle interactions are not interpolated, 
but we instead impose a sharp cutoff on the force
for pairs that air closer than a tenth of a grid cell.

@@

Finally, a choice must be done concerning the longest range of the force. Gravity can be either a) an accurate $1/r^2$ force, as far as the volume allows, 
or b) modified to correctly match the periodicity of the boundary conditions. By default, the code is configured along to the second choice,
which accurately models the growth of structure at very large scales. However, detailed studies of gravitational collapse would benefit 
from the first settings. {\bf (Hugh, are the above two paragraphs correct?)}

>>>>

This section reviews how Poisson's equation is solved on a double-mesh configuration. 
Many parts of the algorithm are identical to {\small PMFAST}, hence we refer the reader 
to section 2 of MPT for more details. In {\small CUBEP3M}, the mass default assignment scheme are
a `cloud-in-cell' (CIC) interpolation for the coarse grid,  and a `nearest-grid-point' (NGP) interpolation 
for the fine grid \citep{1981csup.book.....H}. This choice is motivated by the fact that the most straightforward 
way to implement a p$^3$m algorithm on a mesh is to have exactly zero mesh force inside a grid, 
which is only true for the NGP interpolation. Although CIC generally has a smoother and more accurate force,
the pp implementation enhances the code resolution almost by an order of magnitude. 
We impose a softening length of one tenth of a grid cell to prevent
large scattering; particles seperated by less than this distance have their 
particle-particle force set to zero.

@@

In addition, {\small PMFAST} could run with a different kernel, described in MPT as the `least square kernel', 
which basically adjust the kernel on cell by cell basis based on minimization of the deviation with respect
to the Newtonian predictions. This was originally computed such as to optimize the force calculation for
the case where both grids are computed from CIC interpolation. Moving to a mix CIC/NGP scheme
requires solving the system of equations with the new configuration, a straightforward operation. 
{\bf (results are presented somewhere?)}

Finally, a choice must be done concerning the longest range of the force. Gravity can be either a) an accurate $1/r^2$ force, as far as the volume allows, 
or b) modified to correctly match the periodicity of the boundary conditions. By default, the code is configured along to the second choice,
which accurately models the growth of structure at very large scales. However, detailed studies of gravitational collapse would benefit 
from the first settings.

### 2012-05-31 ScalingPerformance.tex
Incorporated Hughs comments, plus a catastrophic plot showing what happens if we turn off the random offset in the current code

\begin{figure*}%[ht]
%  \vskip -0.5cm 
  \begin{center}
    \includegraphics[width=3.2in]{graphs/scaling_cubep3m_curie.eps}
    \includegraphics[width=3.2in]{graphs/scaling_cubep3m_new.eps}
%  \vskip -1.2cm 
%  \vskip -0.5cm 
  \caption{Scaling of {\small CUBEP3M} on Curie fat nodes (left) and 
    on Ranger TACC facility for very large number of cores (right). Plotted is the code speedup 
    ($N_{\rm particles}^3/t_{\rm wallclock}$) against core count, normalized by the smallest run 
    in each case. Dashed line indicates the ideal weak 
    scaling. The data is listed in Table \ref{summary_scaling_table}.
    \label{scaling}
% \vskip -0.9cm 
}
\end{center}
\end{figure*}

\begin{table*}%[ht]
  \vskip -0.5cm 
  \begin{center}
\caption{Scaling of {\small CUBEP3M} on Curie. Speedup is 
scaled to the smallest run.}
\label{summary_scaling_table}
\begin{tabular}{@{}|llllll|}
\hline
number of cores & speedup & ideal speedup & absolute timing (min) & 
$N_{\rm particles}$& box size ($h^{-1}$Mpc)
\\[2mm]\hline
%\hline
32  &  1.00 & - &3.2 & $256^3$ & 256\\
256  & 7.21 & 8 &3.55 & $512^3$  & 512\\
864  & 25.63 & 27 &4.8 & $864^3$  & 864\\
2048  & 61.87 & 64 &26.48 & $2048^3$ & 2048 \\
\hline
\end{tabular}
\caption{Scaling of  {\small CUBEP3M} for large number of cores 
(on Ranger). Speedup is scaled to the smallest run.}
\label{summary_scaling_table2}
\begin{tabular}{@{}|llllll|}
\hline
number of cores & speedup & ideal speedup & absolute timing (min) & 
$N_{\rm particles}$& box size ($h^{-1}$Mpc)
\\[2mm]\hline
%\hline
864    & 1.00  & -    &258   & $1728^3$  & 6.3\\
4096   & 4.53  & 4.74 &320   & $3072^3$  & 11.4\\
10976  & 9.78  & 12.7 &845   & $5488^3$  & 20\\
21952  & 14.73 & 25.4 &561   & $5488^3$  & 20 \\
\hline
\end{tabular}
\end{center}
  \vskip -0.7cm 
\end{table*}

%the Ranger system at the Texas Supercomputing
%Center (in top 20 in the world) which is a SunBlade x6420 
%with AMD x86\_64 Opteron Quad Core 2300 MHz (9.2 GFlops)
% ``Barcelona'' processors and Infiniband networking. It has 
%a total of 62976 computing cores and 125952 GB of total 
%memory. Its nodes consist of 4 Quad Core processors and 32 GB 
%of shared RAM. For efficiency reasons (local memory access) we 
%typically use smaller MPI 'nodes' consisting of one Quad Core 
%processor and 8 GB of RAM.

@@

@@

@@

{\bf Bottle neck in the largest runs?}

{\bf (Hugh, do you have anything you would like to share 
concerning the largest runs you did on Blue Gene? )}

>>>>

\begin{figure*}%[ht]
%  \vskip -0.5cm 
  \begin{center}
    \includegraphics[width=3.2in]{graphs/scaling_cubep3m_curie.eps}
    \includegraphics[width=3.2in]{graphs/scaling_cubep3m_new.eps}
%  \vskip -1.2cm 
%  \vskip -0.5cm 
  \caption{Scaling of {\small CUBEP3M} on Curie fat nodes (left) and 
    on Ranger TACC facility for very large number of cores (right). Plotted is the code speedup 
    ($N_{\rm particles}^3/t_{\rm wallclock}$) against core count, normalized by the smallest run 
    in each case. Dashed line indicates the ideal weak 
    scaling. The data are listed in Table \ref{summary_scaling_table}.
    \label{scaling}
% \vskip -0.9cm 
}
\end{center}
\end{figure*}

\begin{table*}%[ht]
  \vskip -0.5cm 
  \begin{center}
\caption{Scaling of {\small CUBEP3M} on Curie. Speedup is 
scaled to the smallest run.}
\label{summary_scaling_table}
\begin{tabular}{@{}|llllll|}
\hline
number of cores & speedup & ideal speedup & absolute timing (min) & 
$N_{\rm particles}$& box size ($h^{-1}$Mpc)
\\[2mm]\hline
%\hline
32  &  1.00 & - &3.2 & $256^3$ & 256\\
256  & 7.21 & 8 &3.55 & $512^3$  & 512\\
864  & 25.63 & 27 &4.8 & $864^3$  & 864\\
2048  & 61.87 & 64 &26.48 & $2048^3$ & 2048 \\
\hline
\end{tabular}
\caption{Scaling of  {\small CUBEP3M} on Ranger. Speedup is scaled to the smallest run.}
\label{summary_scaling_table2}
\begin{tabular}{@{}|llllll|}
\hline
number of cores & speedup & ideal speedup & absolute timing (min) & 
$N_{\rm particles}$& box size ($h^{-1}$Mpc)
\\[2mm]\hline
%\hline
864    & 1.00  & -    &258   & $1728^3$  & 6.3\\
4096   & 4.53  & 4.74 &320   & $3072^3$  & 11.4\\
10976  & 9.78  & 12.7 &845   & $5488^3$  & 20\\
21952  & 14.73 & 25.4 &561   & $5488^3$  & 20 \\
\hline
\end{tabular}
\end{center}
  \vskip -0.7cm 
\end{table*}

%the Ranger system at the Texas Supercomputing
%Center (in top 20 in the world) which is a SunBlade x6420 
%with AMD x86\_64 Opteron Quad Core 2300 MHz (9.2 GFlops)
% ``Barcelona'' processors and Infiniband networking. It has 
%a total of 62976 computing cores and 125952 GB of total 
%memory. Its nodes consist of 4 Quad Core processors and 32 GB 
%of shared RAM. For efficiency reasons (local memory access) we 
%typically use smaller MPI 'nodes' consisting of one Quad Core 
%processor and 8 GB of RAM. 
{\bf (This whole section needs to be edited for readability... Plus some things are already in the introduction.)}

@@

Obviously, for maximal efficiency, the number of tiles per node should be a multiple of the number of available processors per {\small MPI} process.
Given the available freedom in the parallel configuration, as long as the load is balanced, it is generally good practice to maximize the number of {\small OPENMP} threads and minimize the number of {\small MPI} processes. 
The information exchange between cores that are part of the same motherboard is generally much faster,
so we minimize the overhead; in addition, having fewer {\small MPI} processes reduces the total amount of buffer zones, a memory that can be used to increase the physical resolution. 
As one scales probes deeper into the non-linear regime however, 
the formation of dense objects can cause memory problems in such configurations, and increasing 
the number of {\small MPI} processes helps to ensure memory locality,
especially in non-uniform memory access (numa) environments.

@@

The intermediary version between {\small PMFAST} and  the current p$^3$m code -- {\small CUBEPM} --
was also ported to the IBM Blue Gene/L platform, and achieved 
weak-scaling up to 4096 processes, with {\small CUBEPM} only incurring a 10 per cent overhead 
at runtime (compared to 8 processes) for a balanced workload.  In order to 
accomdate the limited amount of memory available per processing core on the 
Blue Gene/L platform machines, it was necessary to perform the (long range) MPI FFT
with a volumetric decomposition \citep{3DFFT}.
Slab decomposition would have required a volume too large to fit in system 
memory given the constraints in the simulation geometry.

@@
